---
title: "PA1_template"
author: "H20Lady"
date: "Saturday, October 11, 2014"
output: html_document

---

##Assignment
****

```{r, echo=TRUE}

#set working directory in R to folder with database files
act <- read.csv("activity.csv")
head(act)
# Format Date column in R
act$dateconverted <- as.Date(act$date, format="%Y-%m-%d")

# Create a data frame grouped by Day column and Summed by Steps column

library(dplyr)

act1 <- 
     act %>%
     group_by(dateconverted) %>%
     summarise(totsteps = sum(steps, na.rm=TRUE))

```

****
###What is mean total number of steps taken per day?
#####Histogram of the total number of steps taken each day. 
######Missing values have been ignored (i.e. removed) in this dataset.
****
```{r, echo=TRUE}

library(ggplot2)

p <- ggplot(act1, aes(x=totsteps)) + geom_histogram(binwidth=1000)

hist<- p + 
     labs(title = "Histogram of Total Number of Steps Per Day\n Missing Values Included") + 
     xlab("Total Number of Steps Per Day") + ylab("Frequency of Total Number of Steps Per Day") + 
     xlim(0,25000) +
     ylim(0, 20)

print(hist)

#Saves as image file to drive:
#ggsave("Fig1_act.png", plot=hist, scale = 1, width = 8, height = 5, units = c("in"), dpi = 80)

```

****
#####Mean and Median of Total number of steps taken per day. 
######Missing values have been ignored in this dataset.
****
```{r,echo=TRUE}
mean(act1$totsteps)
median(act1$totsteps)
```
Mean of Total Steps Per Day = 9354 and Median of Total Steps Per Day = 10395

****
###What is the average daily activity pattern?
#####Time-Series Plot
****
```{r,echo=TRUE}

act2 <- 
     act %>%
     group_by(interval) %>%
     summarise(avgsteps = mean(steps,na.rm=TRUE))


q <- ggplot(act2, aes(x=interval, y=avgsteps)) + geom_line(colour = "red", size = 0.5)

hist1<- q + 
     labs(title = "Time Series Plot of 5 minute-intervals \n From October 1, 2012 to November 30, 2012 \n Missing Values Ignored")+ 
     xlab("Intervals") + ylab("Avg. Number of Steps") + 
     xlim(0,3000) +
     ylim(0, 300)

print(hist1)

#Saves as image file to drive:
#ggsave("Fig2_act.png", plot=hist, scale = 1, width = 8, height = 5, units = c("in"), dpi = 80)

which.max(act2$avgsteps)

act2[104,]
```

The 103rd row's 5-minute interval which is the 835th minute contains the maximum average number of steps

****
###Imputing missing values
#####Histogram of the total number of steps taken each day. 
######This dataset has missing values filled.
****
```{r, echo=TRUE}
summary(act)
```

There are 2304 NAs in the dataset

****
######This dataset has missing values filled.
****
```{r,echo=TRUE}
#find NA values for each day and replace with mean of that day



```
Mean = b and Median = m
****
```{r,echo=TRUE}

act2 <- 
     act %>%
     group_by(dateconverted) %>%
     summarise(totsteps = sum(steps, na.rm=TRUE))

p <- ggplot(act2, aes(x=totsteps)) + geom_histogram(binwidth=1000)

hist<- p + 
     labs(title = "Histogram of Total Number of Steps Per Day\n Missing Values Filled") + 
     xlab("Total Number of Steps Per Day") + ylab("Frequency of Total Number of Steps Per Day") + 
     xlim(0,25000) +
     ylim(0, 20)

print(hist)

#Saves as image file to drive:
#ggsave("Fig3_act.png", plot=hist, scale = 1, width = 8, height = 5, units = c("in"), dpi = 80)

```

###Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}
#find NA values for each day and replace with mean of that day



```



